### 2023-11-10 main.tex
Update main.tex

\vspace{.5cm}
\noindent
\begin{remark}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
(those of types $\{1,\ldots,N-1\}$), we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section \ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.  
\end{remark}

@@

\subsection{Reversible measures (equilibrium)}

@@

with marginals $\mu_{rev}^{x}$ given by 

@@

\subsection{Lie algebraic description}

@@

\section{Duality}\label{sectionDuality}

@@

@@

@@

{\color{blue}Think about replacing $\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|$ with $\langle \xi_{1},\ldots,\xi_{N}|_{u(x)}$}\\

@@

since
\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}=0\dt

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Subsection \ref{subsection-description-process-LINE}. We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. Now we assume the hard-core exclusion rule, i.e. $\nu=1$. The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of $gl(N)$. Duality result stated in Subsection \ref{statementDualitySubsection} can be adapted. The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Sections \ref{sectionIntegrabiliy}.\\

%In this notation, the elements of the duality matrix reads
%\begin{equation}\label{ElementsDualityMatrixChain_1}
%	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
%\\
Using the above duality elements, we can compute the  $m$-point correlations between non-empty particles using the absorption probabilities. Therefore, to determine the non-equilibrium steady state correlations of the integrable multi-species stirring process, it is enough to compute these absorption probabilities.  %For arbitrary $m\in \mathbb{N}$ such that $m<L$, we fix $a_{k}\in\{1,\ldots,N-1\}$ and $x_{k}\in\{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$ for every $k\in\{1,\ldots,m\}$. 
\begin{proposition}[Correlations via duality]
Consider $m\in \{1,\ldots,L\}$ and let $a_{k}\in\{1,\ldots,N-1\}$ species positioned at sites $x_{k}\in\{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$. Then, the $m$-point correlations in the non-equilibrium steady state (NESS) read
\begin{equation}
\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\mathcal{P}(t_{1},\ldots,t_{m},\bm{\xi})
\end{equation}
where
\begin{equation}\label{absProbabilitiesIntegrable}
	\mathcal{P}(t_{1},\ldots,t_{m},\bm{\xi}):=\mathbb{P}\left(\bm{\xi}(\infty)=\sum_{k=1}^{m}\left(t_{k}\delta^{0}_{a_k}+(1-t_{k})\delta^{L+1}_{a_k}\right)\lvert \bm{\xi}(0)=\bm{\xi}\right)
\end{equation}
is the  probability that  $t_{k}$ particles of species $a_{k}$ (initially positioned at site $x_{k}$) are absorbed at the extra site $\{0\}$. The dual configuration is $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, meaning that 
\begin{equation}
	\begin{cases}
		\xi_{a}^{x}=1\qquad \text{if}\;x=x_{k}\;\text{and}\; a=a_{k}\\
		\xi_{a}^{x}=0\qquad \text{otherwise}
	\end{cases}
\end{equation}
\end{proposition}
\textbf{Proof}:
For any $\bm{\tau}\in \Omega$ the elements of the duality matrix defined in \eqref{ElementsDualityMatrixChain_1} read
\begin{equation}
	D(\bm{\tau},\bm{\xi})%=\prod_{k=1}^{m}\mathbbm{1}_{\{\delta_{\tau_{x_{k}},a_{k}}=1\}}
	=\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\dt
\end{equation}
Therefore, by duality we have 
\begin{equation}\label{CorrelationByABS-prob}
	\begin{split}
\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]&=\lim_{t\to \infty}\mathbb{E}_{\bm{\tau}}\left[D(\bm{\tau}(t),\bm{{\xi}})\right]=\lim_{t\to \infty}\mathbb{E}_{\bm{\xi}}\left[D(\bm{\tau},\bm{{\xi}}(t))\right]
	\\&=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\mathcal{P}(t_{1},\ldots,t_{m},\bm{\xi})
\end{split}
\end{equation}
\begin{flushright}
	$\square$
\end{flushright}

\section{Exact steady state and correlations for the integrable chain}\label{sectionIntegrabiliy}
In this section we aim to use duality and the matrix product ansatz (MPA) to determine the moments and correlations of the non-equilibrium steady state for the integrable version of the multi-species stirring process, defined in Subsection \ref{integrableChain-duality}. As it will turn out, the commutation relations of the matrix product ansatz allows to determine a closed formula for all correlations. As a by-product, absorption probabilities can be explicitly found.%, giving a further probabilistic interpretation for the non-equilibrium steady state. 
\subsection{Matrix product ansatz for the non-equilibrium steady state}
The matrix product ansatz for the multi-species stirring process has been formulated in \cite{vanicat2017exact}, here we briefly recall the main steps.
 Denoting by $|\Psi(t)\rangle$ the column vector that encodes the probability distribution of the chain with Hamiltonian \eqref{hamiltonian} at time $t\geq 0$, its evolution equation is given by 
 the master equation
    \frac{d|{\Psi}(t)\rangle}{dt}=H|{\Psi}(t)\rangle\dt
This Markov chain is irreducible and positive recurrent, therefore there exists a unique stationary measure, that will be reached when time goes to infinity, regardless of the initial configuration. 
We denote by $|\Psi\rangle$ the column vector that gives the stationary distribution. This vector is the right eigenvector with vanishing eigenvalue of $H$, i.e. it solves 
\begin{equation}\label{definition-SteadyStateH}
	H|\Psi\rangle =0\dt
The MPA states the following
	|\Psi\rangle=\frac{1}{Z_{L}}\langle V^{*}|\begin{pmatrix}
	\end{pmatrix}\otimes \ldots\otimes \begin{pmatrix}
	\end{pmatrix}|V\rangle
with the normalization 
	Z_{L}=\langle V^{*}|(X_{1}+\ldots +X_{N})^{L}|V\rangle\dt
Here the operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfil the commutators
\begin{equation}\label{bulk}
	\left[X_{a},X_{b}\right]=(\alpha_{a}-\beta_{a})X_{b}-\lambda_{b}X_{a}\qquad\forall a,b=1,\ldots,N
and their action on the boundary vectors are
\begin{equation}\label{leftBoundary}
	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=(\alpha_{a}-\beta_{a})\langle V^{*}|\qquad\forall a=1,\ldots,N\co
\begin{equation}\label{rightBoundary}
	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-(\alpha_{a}-\beta_{a})|V\rangle\qquad\forall a=1,\ldots,N\dt
Without loss of generality, we further assume that $\langle V^{*}|V\rangle=1$. 
\begin{remark} Only $N-1$ of the $N$ equations \eqref{leftBoundary}  are independent. This can be seen by summing them over the species $a\in \{1,\ldots,N\}$, i.e.
		\sum_{a=1}^{N}	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=0,%\sum_{a=1}^{N}(\alpha_{a}-\beta_{a})\langle V^{*}|=0
	and using  \eqref{ratesConditions}.
 Similarly for the right boundary  \eqref{rightBoundary} there are only  $N-1$ independent equations. \end{remark} 
The MPA gives an abstract form of the non-equilibrium steady state in terms of an algebra of operators. However, the computation of exact correlations are in general involving. 
 We will show that the existence of absorbing dual process established in Section \ref{sectionDuality} simplifies substantially the use of commutators in the
 computation. The idea is summarized as follows. We define a sequence of similarity transformation such that 

@@

\subsubsection{Explicit examples of non-equilibrium steady state correlations}
We give examples  of correlations for up to three points applying formula \eqref{exactCorrelations}.

@@

	\mathcal{P}(0,\bm{\xi})=1-\frac{(L+1-x_{1})}{(1+L)}\qquad \mathcal{P}(1,\bm{\xi})=\frac{(L+1-x_{1})}{(1+L)}

>>>>

\noindent
\begin{remark}
\label{hole}
In the following we will often consider type $N$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types. It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
(those of types $\{1,\ldots,N-1\}$), we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
The choice of type $N$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section \ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $N$ on the graph.
Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.  
\end{remark}

@@

\subsection{Reversible measures in the equilibrium set-up}

@@

with marginals $\mu_{\text{rev}}^{x}$ given by 

@@

\subsection{Lie algebraic description of the process}

@@

\section{Duality}\label{sectionDuality}
In this section we prove the first main result of this paper. We show that the multispecies stirring process with open boundary
can be studied by means of a dual process. After recalling the definition of duality between two
Markov processes in Section \ref{def-duality}, we formulate duality  in Theorem \ref{thm-duality}. A proof of the theorem is given in Section \ref{proof-th-duality}.
The duality result is important since it brings in several simplifications, the main one being that the
dual process has absorbing sites. Therefore, as shown in Proposition \ref{prop-absorbing}, the study of the non-equilibrium steady state for  the multispecies stirring process with open boundary
is reduced to the study of the absorption probabilities of dual particles.

@@

\label{def-duality}

@@

\label{proof-th-duality}

@@

%{\color{blue}Think about replacing $\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|$ with $\langle \xi_{1},\ldots,\xi_{N}|_{u(x)}$}\\

@@

%since
%\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N}^{u(x)}|(\mathbf{a}^{+})_{a}^{u(x)}=0\dt

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Subsection \ref{subsection-description-process-LINE} with the hard-core constrain, i.e. $\nu=1$. 
We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. 
We also recall the notation for the occupation variable which is called $\bm{\tau}=(\tau_1,\ldots,\tau_L)$ with $\tau_x\in\{1,\ldots,N\}$.
The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of $gl(N)$. The duality result stated in Subsection \ref{statementDualitySubsection} can be adapted. The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Section \ref{sectionIntegrabiliy}.\\

%In this notation, the elements of the duality matrix reads
%\begin{equation}\label{ElementsDualityMatrixChain_1}
%	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
%\\
Using duality, one can compute the  $m$-point correlations between non-empty particles in terms of the absorption probabilities of $m$ dual particles. Therefore, to determine the non-equilibrium steady state correlations of the integrable multi-species stirring process, it is enough to compute these absorption probabilities.  %For arbitrary $m\in \mathbb{N}$ such that $m<L$, we fix $a_{k}\in\{1,\ldots,N-1\}$ and $x_{k}\in\{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$ for every $k\in\{1,\ldots,m\}$. 
\begin{proposition}[Correlations via duality]
\label{prop-absorbing}
Consider $m\in \{1,\ldots,L\}$ and let $a_{k}\in\{1,\ldots,N-1\}$ species positioned at sites $x_{k}\in\{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$. Then, the $m$-point correlations in the non-equilibrium steady state (NESS) read
\begin{equation}
\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\mathcal{P}_{x_1,\ldots,x_m}(t_{1},\ldots,t_{m})
\end{equation}
where
\begin{equation}\label{absProbabilitiesIntegrable}
	\mathcal{P}_{x_1,\ldots,x_m}(t_{1},\ldots,t_{m}):=\mathbb{P}\left(\bm{\xi}(\infty)=\sum_{k=1}^{m}\left(t_{k}\delta^{0}_{a_k}+(1-t_{k})\delta^{L+1}_{a_k}\right)\lvert \bm{\xi}(0)=\bm{\xi}\right)
\end{equation}
is the  probability that  $t_{k}$ dual particles of species $a_{k}$ (initially positioned at site $x_{k}$) are absorbed at the extra site $\{0\}$. The dual initial configuration is $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, meaning that 
\begin{equation}
\label{xsi-init}
	\begin{cases}
		\xi_{a}^{x}=1\qquad \text{if}\;x=x_{k}\;\text{and}\; a=a_{k}\\
		\xi_{a}^{x}=0\qquad \text{otherwise}
	\end{cases}
\end{equation}
\end{proposition}
\textbf{Proof}:
For any $\bm{\tau}\in \Omega$ the duality matrix defined in \eqref{ElementsDualityMatrixChain_1} evaluated on the dual configuration $\bm{\xi}\in\widetilde{\Omega}$ given in \eqref{xsi-init} read
\begin{equation}
	D(\bm{\tau},\bm{\xi})%=\prod_{k=1}^{m}\mathbbm{1}_{\{\delta_{\tau_{x_{k}},a_{k}}=1\}}
	=\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\dt
\end{equation}
Therefore, by duality and ergodicity we have 
\begin{equation}\label{CorrelationByABS-prob}
	\begin{split}
\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]&=\lim_{t\to \infty}\mathbb{E}_{\bm{\tau}}\left[D(\bm{\tau}(t),\bm{{\xi}})\right]=\lim_{t\to \infty}\mathbb{E}_{\bm{\xi}}\left[D(\bm{\tau},\bm{{\xi}}(t))\right]
	\\&=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\mathcal{P}(t_{1},\ldots,t_{m},\bm{\xi})\dt
\end{split}
\end{equation}
\begin{flushright}
	$\square$
\end{flushright}

\section{Exact steady state and correlations for the integrable chain}\label{sectionIntegrabiliy}
In this section we combine duality and the matrix product ansatz (MPA) to determine the correlations of the non-equilibrium steady state for the integrable version of the multi-species stirring process, defined in Subsection \ref{integrableChain-duality}. As it will turn out, such combination of duality and MPA allows to determine a closed formula for the absorption correlations and thus (using Proposition \ref{prop-absorbing}) for the correlations.
\subsection{Matrix product ansatz for the non-equilibrium steady state}
The matrix product ansatz for the multi-species stirring process has been formulated in \cite{vanicat2017exact}, here we briefly recall the main steps.
 Denoting by $|\Psi(t)\rangle$ the column vector that encodes the probability distribution of the chain with Hamiltonian \eqref{hamiltonian} at time $t\geq 0$, its evolution equation is given by 
 the master equation
\begin{equation}
    \frac{d|{\Psi}(t)\rangle}{dt}=H|{\Psi}(t)\rangle\dt
\end{equation}
This Markov chain is irreducible and positive recurrent, therefore there exists a unique stationary measure, that will be reached when time goes to infinity, regardless of the initial configuration. 
We denote by $|\Psi\rangle$ the column vector that gives the stationary distribution (non-equilibrium steady state). This vector is the right eigenvector with vanishing eigenvalue of $H$, i.e. it solves 
\begin{equation}\label{definition-SteadyStateH}
	H|\Psi\rangle =0\dt
\end{equation}
The MPA states the following
\begin{equation}
	|\Psi\rangle=\frac{1}{Z_{L}}\langle V^{*}|\begin{pmatrix}
		X_{1}\\
		\vdots\\
		X_{N}
	\end{pmatrix}\otimes \ldots\otimes \begin{pmatrix}
		X_{1}\\
		\vdots\\
		X_{N}
	\end{pmatrix}|V\rangle
\end{equation}
with the normalization 
\begin{equation}
	Z_{L}=\langle V^{*}|(X_{1}+\ldots +X_{N})^{L}|V\rangle\dt
\end{equation}
Here the operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfil the commutators
\begin{equation}\label{bulk}
	\left[X_{a},X_{b}\right]=(\alpha_{a}-\beta_{a})X_{b}-(\alpha_{b}-\beta_{b})X_{a}\qquad\forall a,b=1,\ldots,N
\end{equation}
and their action on the boundary vectors are
\begin{equation}\label{leftBoundary}
	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=(\alpha_{a}-\beta_{a})\langle V^{*}|\qquad\forall a=1,\ldots,N\co
\end{equation}
\begin{equation}\label{rightBoundary}
	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-(\alpha_{a}-\beta_{a})|V\rangle\qquad\forall a=1,\ldots,N\dt
\end{equation}
Without loss of generality, we further assume that $\langle V^{*}|V\rangle=1$. 
\begin{remark} Only $N-1$ of the $N$ equations \eqref{leftBoundary}  are independent. This can be seen by summing them over the species $a\in \{1,\ldots,N\}$, i.e.
	\begin{equation}
		\sum_{a=1}^{N}	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=0,%\sum_{a=1}^{N}(\alpha_{a}-\beta_{a})\langle V^{*}|=0
	\end{equation}
	and using  \eqref{ratesConditions}.
 Similarly for the right boundary  \eqref{rightBoundary} there are only  $N-1$ independent equations. \end{remark} 
The MPA gives an abstract form of the non-equilibrium steady state in terms of an algebra of operators. However, the computation of  correlations are in general involved.

 In the next Section \ref{subsection-exact} we will show that the existence of absorbing dual process established in Section \ref{sectionDuality} simplifies substantially the algebra of operators. The idea is summarized in the following scheme. 
 We define a sequence of similarity transformation such that 

@@

\subsubsection{Examples of non-equilibrium steady state correlations}
We give examples  of correlations for $m=1,2,3$ applying formula \eqref{exactCorrelations}.

@@

	\mathcal{P}(0)=1-\frac{(L+1-x_{1})}{(1+L)}\qquad \mathcal{P}(1)=\frac{(L+1-x_{1})}{(1+L)}
